Data analysis for employee cluster descriptive statistics

### 2023-06-07 04. Analysis/scripts/03-02-data-analysis-employee-cluster-assessing-clusters.Rmd
Initial commit
Mainly worked on decriptive statistics. Other file is to review the different numbers and types of clusters.

---
title: "Data analysis for employee categories"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "prep_employees_2023-05-24.rdata"
  unique_identifier_1: !r c("date","person_id")
  date_cohort_month: "2022-03-01"
  observation_window_length_months: 6
  in_number_of_clusters_start: 1
  in_number_of_clusters_end: 10

@@

**Unique identifier for data file 1**: `r params$unique_identifier_1`

**Cohort start month**: `r nice_month_year(date_cohort_month_start)`

**Observation window**: `r nice_month_year(date_cohort_month_start)` to `r nice_month_year(date_observation_window_month_end)` (`r params$observation_window_length_months` months)

**Number of clusters**: Testing `r params$in_number_of_clusters_start` to `r params$in_number_of_clusters_end`

@@

```{r}
prep_employees %>% 
  skim()
```

@@

Add underscore to specification.

```{r}
kclusts <- 
  kclusts %>% 
  mutate(
    specification = str_replace_all(specification," ","_")
  )
```

Check `specification` results.

```{r}
kclusts %>% 
  count(specification)
```

>>>>

---
title: "Data analysis for employee cluster descriptive statistics"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "clusters-employee-2023-06-01.rdata"
  in_number_of_clusters: 5
  in_specification: "scale_working_proportion_pairs_1_10"

@@

**Number of clusters**: `r params$in_number_of_clusters`

**Specification** `r params$in_specification`

@@

## Select and unnest data {.tabset .tabset-pills}

### Filter to specification and cluster

Filter to data with `r params$in_number_of_clusters` clusters with specification `r params$in_specification`.

```{r}
clust <- 
  kclusts %>% 
  filter(
    k == params$in_number_of_clusters &
      specification == params$in_specification
  )
```

### Unnest

Filter to data with `r params$in_number_of_clusters` clusters with specification `r params$in_specification`.

```{r}
clusters <- 
  kclusts %>%
  unnest(cols = c(tidied))

assignments <- 
  kclusts %>% 
  unnest(cols = c(augmented))

clusterings <- 
  kclusts %>%
  unnest(cols = c(glanced))

```

@@

### 2023-06-07 04. Analysis/scripts/03-03-data-analysis-employee-cluster-descriptive-statistics.Rmd
Initial commit
Mainly worked on decriptive statistics. Other file is to review the different numbers and types of clusters.

---
title: "Data analysis for employee cluster descriptive statistics"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "clusters-employee-2023-06-01.rdata"
  in_number_of_clusters: 5
  in_specification: "scale_working_proportion_pairs_1_10"

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
library(tidymodels)
```

```{r cohort dates}
date_cohort_month_start <- ymd(params$date_cohort_month)
date_cohort_month_end <- ceiling_date(date_cohort_month_start,unit = "month") %>% rollbackward()
date_observation_window_month_start <- date_cohort_month_start %m+% months(params$observation_window_length_months)
date_observation_window_month_end <- ceiling_date(date_observation_window_month_start,unit = "month") %>% rollbackward()
```

```{r}
set.seed(27)
```

@@

@@

@@

```{r}
clusters <- 
  kclusts %>%
  unnest(cols = c(tidied))

assignments <- 
  kclusts %>% 
  unnest(cols = c(augmented))

clusterings <- 
  kclusts %>%
  unnest(cols = c(glanced))

@@

>>>>

---
title: "Data analysis for employee cluster descriptive statistics"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "clusters-employee-2023-06-01.rdata"
  data_file_2: "prep_employees_2023-05-24.rdata"
  in_number_of_clusters: 5
  in_specification: "scale_working_proportion_pairs_1_10"

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("geomtextpath" %in% installed.packages()[,"Package"]))
{
  install.packages("geomtextpath")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
library(tidymodels)
```

```{r cohort dates}
# date_cohort_month_start <- ymd(params$date_cohort_month)
# date_cohort_month_end <- ceiling_date(date_cohort_month_start,unit = "month") %>% rollbackward()
# date_observation_window_month_start <- date_cohort_month_start %m+% months(params$observation_window_length_months)
# date_observation_window_month_end <- ceiling_date(date_observation_window_month_start,unit = "month") %>% rollbackward()
```

@@

**Data file 2**: `r params$data_file_2`

@@

Load `r params$data_file_2` from `r params$data_folder_1` data folder.

```{r}
load(file = file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1,params$data_file_2))
```

```{r}
prep_employees %>% 
  skim()
```

@@

```{r}
clusters <- 
  clust %>%
  unnest(cols = c(tidied))

assignments <- 
  clust %>% 
  unnest(cols = c(augmented))

clusterings <- 
  clust %>%
  unnest(cols = c(glanced))

@@

```{r}
assignments %>% 
  glimpse()
```

### Join assignments to employee data

```{r}
d <-
  assignments %>%
  select(k, person_id, .cluster, specification) %>%
  full_join(prep_employees) %>% 
  mutate(
    .cluster = fct_na_value_to_level(.cluster, level = "(Missing)")
  )
```
Review data.

## Work pattern statistics {.tabset .tabset-pills}

```{r}
summaries <- function(in_data,...)
{
  in_data %>% 
  group_by(...) %>% 
  summarise(
    across(
      .cols = c(daily_status_teleworking,daily_status_reporting_to_job_site,
                daily_status_teleworking_working_prop,daily_status_reporting_to_job_site_working_prop,
                daily_status_teleworking_observation_prop,daily_status_reporting_to_job_site_observation_prop,
                starts_with("mon_daily_"),
                starts_with("tue_daily_"),
                starts_with("wed_daily_"),
                starts_with("thu_daily_"),
                starts_with("fri_daily_"),
                ends_with("_cor_telework"),
                ends_with("_cor_status"),
                ends_with("prop_pairs"),
                team_size_mode,
                starts_with("days_colocated"),
                starts_with("days_telework"),
                working_days
                ),
      .fns = ~mean(.,na.rm = TRUE)
    )
  ) %>% 
    mutate(
      statistic = "mean"
    )
}

filter_rule_base <- function(in_summaries_data,in_filter)
{
  filt_var_mod <- enquo(in_filter) 
  in_summaries_data %>% 
     dplyr::filter((!!filt_var_mod) == 1) %>% 
    pivot_longer(cols = !!filt_var_mod) %>% 
    mutate(
      temp = str_remove(name,"rule_based_"),
      temp = str_replace_all(temp,"_"," "),
      temp = str_to_sentence(temp),
      group = glue::glue("Rule-based: {temp}")
    ) %>% 
    select(-name, -value,-temp)
}

summaries_rule_based <- function(in_data,in_rule_based_var)
{
  d %>% 
    summaries({{in_rule_based_var}}) %>% 
    filter_rule_base({{in_rule_based_var}})
}

prep_a <- bind_rows(
  d %>%
    summaries() %>%
    mutate(group = "Whole cohort"),
  d %>%
    summaries(.cluster) %>%
    mutate(group = glue::glue("Cluster: {.cluster}")) %>%
    select(-.cluster),
  d %>% 
    summaries_rule_based(rule_based_all_remote),
  d %>% 
    summaries_rule_based(rule_based_primarily_teleworking_observation),
  d %>% 
    summaries_rule_based(rule_based_primarily_teleworking_working),
  d %>% 
    summaries_rule_based(rule_based_primarily_reporting_to_job_site_observation),
  d %>% 
    summaries_rule_based(rule_based_primarily_reporting_to_job_site_working),
  d %>% 
    summaries_rule_based(rule_based_mostly_teleworking_observation),
  d %>% 
    summaries_rule_based(rule_based_mostly_teleworking_working),
  d %>% 
    summaries_rule_based(rule_based_mostly_reporting_to_job_site_observation),
  d %>% 
    summaries_rule_based(rule_based_mostly_reporting_to_job_site_working),
  d %>% 
    summaries_rule_based(rule_based_position_category_missing),
  d %>% 
    summaries_rule_based(rule_based_position_category_onsite_flexible),
  d %>% 
    summaries_rule_based(rule_based_position_category_offsite),
  d %>% 
    summaries_rule_based(rule_based_position_category_onsite_required),
  d %>% 
    summaries_rule_based(rule_based_position_category_position_has_changed),
  d %>% 
    summaries_rule_based(rule_based_position_category_not_specified),
  d %>% 
    summaries_rule_based(rule_based_telework_onsite_minimum),
)
```

### Daily telework

```{r}
prep_a %>% 
  select(group,statistic,starts_with("daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Daily telework by day of week

#### Monday

```{r}
prep_a %>% 
  select(group,statistic,starts_with("mon_daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

#### Tuesday

```{r}
prep_a %>% 
  select(group,statistic,starts_with("tue_daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

#### Wednesday

```{r}
prep_a %>% 
  select(group,statistic,starts_with("wed_daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

#### Thursday

```{r}
prep_a %>% 
  select(group,statistic,starts_with("thu_daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

#### Friday

```{r}
prep_a %>% 
  select(group,statistic,starts_with("fri_daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Daily reporting to site

```{r}
prep_a %>% 
  select(group,statistic,starts_with("daily_status_reporting_to_job_site")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Team size

```{r}
prep_a %>% 
  select(group,statistic,starts_with("team_size")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Working days

```{r}
prep_a %>% 
  select(group,statistic,starts_with("working_days")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Co-located days with at least one team member

```{r}
prep_a %>% 
  select(group,statistic,starts_with("days_colocated_1")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Co-located days with at whole team

```{r}
prep_a %>% 
  select(group,statistic,starts_with("days_colocated_team_")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Co-located days with supervisor

```{r}
prep_a %>% 
  select(group,statistic,starts_with("days_colocated_supervisor_")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Days team teleworks

```{r}
prep_a %>% 
  select(group,statistic,starts_with("days_telework_team_")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

The principle work pattern statistics are as follows:

-   Employee work pattern statistics:

    -   Proportion in employee time onsite by the day of the week

    -   Correlations of by employee time between the days of the week

    -   Number exiting employment

    -   Number changing supervisors

## Descriptive statistics {.tabset .tabset-pills}

 Employee descriptive statistics

    -   Number and proportion of employees by grade

    -   Number and proportion of employees by occupational series

    -   Number and proportion of employees by position category

    -   Number and proportion of employees by GSA region

    -   Number and proportion of employees by SSO

    -   Number and proportion of employees by state

    -   Number and proportion of employees by city

    -   Number and proportion of employees by supervisor level

## Distributions by cluster {.tabset .tabset-pills}

### Function

Function to create a density plot for a specific numeric variable.

```{r}
gg_density <- function(in_data,in_variable)
{
  in_variable_sym <- sym(in_variable)

  spec <- unique(in_data$specification)

  clus <- unique(in_data$k)

  in_data %>% 
  ggplot(aes(x = !!in_variable_sym, color = .cluster, fill = .cluster)) +
  geom_density(alpha = .25) +
    geomtextpath::geom_textdensity(aes(label = .cluster), 
                   hjust = "ymax", 
                   vjust = 1, text_only = TRUE, text_smoothing = 0, straight = TRUE) + 
    oescolorrrs::scale_color_oes(palette = 'accent colors') +
    oescolorrrs::scale_fill_oes(palette = 'accent colors') +
    theme_fivethirtyeight() +
    labs(
      title = glue::glue("Histogram of `{in_variable}`"),
      caption = glue::glue("For specification {spec} with {clus} clusters.")
    ) +
    theme(
      legend.position = "none",
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "white"),
      # panel.border = element_blank()
    )
}
```

### List of numeric variables

```{r}
num_vars <- 
  d %>% 
  select_if(is.numeric) %>% 
  select(-k) %>% 
  names()

num_vars
```

### Distributions

```{r}
num_vars[1:20] %>% 
  map(~gg_density(d,in_variable = .))
```

## Descriptive

The descriptive statistics are:

-  
The employees statistics will be calculated for the following groups:

1.  Whole cohort of employees
2.  By rule-based categories of work pattern
3.  By cluster defined categories of work pattern
